Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-01-04 Ben_Manuscripts/structure_paper/Draft.tex
Minor edits

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanoporous
  membranes with high selectivity permitted by a precisely controlled pore size
  and high permeability allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
  are suitable candidates for aqueous separation applications. LLCs are
  amphiphilic molecules that share the characteristic ability of nanostructured
  porous membrane materials to create highly ordered nanostructures with the
  added benefits of low cost and synthetic techniques feasible for large-scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
  wt\% added water results in formation of the lyotropic H\textsubscript{II}
  phase. In both cases, Na-GA3C11 monomers assemble into mesophases made of
  hexagonally packed, uniform size, cylinders with hydrophilic head groups
  oriented inward towards the cylinder center. This LLC assembly is then
  polymerized in situ to form a cross-linked polymer network to stabilize the
  structure. The hydrophilic region can act as a pore for aqueous separations
  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
  specific separations by changing the monomer chemistry
  \cite{resel_h2-phase_2000}.

  Research into polymerized LLC membranes has been revived in recent years.
  During early stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  one can align the mesophases using a magnetic field with subsequent
  crosslinking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications.  Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}.  In a later study of rejection performance, it
  was shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} cannot perform separations of solutes less than 1.2 nm in
  diameter because the pores are too large \cite{zhou_supported_2005}.  We do not
  yet understand how to controllably reduce the effective pore size or how to
  tune the chemical environment in the nanopores of this or related materials for
  effective water desalination and small organic molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. It will be challenging to efficiently narrow down
  the large design space in a laboratory setting without a robust model.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the volume of the liquid crystal monomers
  proposes that each layer consists of seven monomers
  \cite{resel_structural_2000}.  A molecular model orders of magnitude larger
  than any other reported atomistic liquid crystal membrane simulations, as we
  present here, has the best chance of directly answering this question.  We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations versus a sandwiched
  conformation \cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006} In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configuration is most likely.

@@

	\item Within R-alkanes, are four spots of higher relative intensity.
	Accordingly, we name these reflection R-spots. Thin location of all spots is $\sim 37$
	degrees from the $q_z$ axis in their respective quadrants. In many liquid
	crystal systems one can explain the spots by the tilt angle of the alkane chains
	with respect to the membrane plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane. This plane is geometrically related the distance
	between pores.  The feature, which we named R-pores, is characterized by dots
	along the equatorial axis defined when $q_z$ = 0. The spacing between dots is
	indicative of the hexagonal symmetry of the packed pores. We observe the same
	information at higher resolution using SAXS (Fig.~\ref{fig:SAXS}).

@@

  We constructed an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000K to 50K over 10
  nanoseconds. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration the trajectory which we used for full system
  construction (Figure~\ref{fig:python}a). See supplemental information for
  further detail.

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane. We explore two different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial layer spacing of 5
  \AA. We briefly explored a third system with an initial layer spacing of 10
  \AA. If we initially space layers too far apart, they will collapse on each
  other while simultaneously slipping in between layers of adjacent pores, which
  leads to an artificially thick membrane with pores spaced closely together.
  Further details of simulations with large layer spacing are in the supplemental
  information.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, the we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

@@

  \subsection{Crosslinking}

  In order to fully match synthetic procedures, we created a crosslinking
  algorithm that one can apply to equilibrated structures. The purpose of
  crosslinking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data. We
  developed the algorithm based on the known reaction mechanism. Crosslinking of
  this system is a free radical polymerization (FRP) taking place between
  terminal vinyl groups present on each of the three monomer tails. FRPs require
  an initiator which bonds to the system, meaning new atoms are introduced into
  the system. For simplicity, we simulated the initiator as hydrogen and made it
  present in the simulation by including them in all possible locations where an
  addition could occur as dummy atoms. We carry out the crosslinking procedure
  iteratively. During each iteration, the algoritm selects eligible bonding
  carbon atoms based on a distance cut-off. The topology is updated with new
  bonds and dummy hydrogen atoms are changed to appropriate hydrogen types.
  Head-to-tail addition was the only propagation mode considered due to its
  dominance in the real system \cite{young_introduction_2011}. We did not
  consider direction of attack because the resultant mixture is racemic.

  Our implementation requires long simulation times to achieve high crosslink 
  densities. A typical crosslinking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a crosslinked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not crosslink all systems tested, but only
  the most promising structure. We show that crosslinking does not significantly
  change any of our drawn conclusions in Section 3.6.

@@

  We normalized the colorbars on all diffraction patterns relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$.  We exclude intensities
  within $\pm$ 30\degree~of the $q_r=0$ axis, since the simulated patterns differ
  from experiment in those regions.  Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.

@@

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation relates the DC ionic conductivity to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$ 
  Sodium ion diffusion coefficients were found by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell.

@@

%MRS4: can we come up with another name?  I still think disordered pores is a decent name, since there is no pi-stacking layering.  ``structures that form when layers are 5 AA apart'' is not so good.  or ``more disordered''?
%BJC3: agreed, I'll stick with disordered pores
%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom~as suggested by experiment.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). For reasons to become
  aparent later, we will refer to this class of systems as the disordered pore
  basin.  Systems which we build with layers initially stacked 3.7 \AA~apart will
  be referred to as the ordered basin. We studied each basin in both the parallel
  displaced and sandwiched configurations.

@@

  In all cases, the space in the pore region is filled with sodium ions and
  head groups. In ordered pore systems, there is less density in the center of
  the pore indicating a more ordered pore structure. In both the sandwich and
  parallel displaced configurations, we see the density of head groups and sodium
  ions fall to less than 50\% of its maximum at r = 0
  (Fig.~\ref{fig:layered_density}). The situation is most pronounced in the
  sandwiched configuration where the maximum head group density occurs 0.44 nm
  from the pore center. The parallel displaced configuration reaches its maximum
  0.35 nm from the pore center. In contrast, both disordered pore systems show
  very little difference in density from its maximum. This implies a more uniform
  distribution of head groups within the pore center. 
  % High vacancy is less likely to be stable in the real system

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the crosslinking algorithm to the equilibrated
  parallel-displaced configuration built with 5 monomers per layer. We simulated
  the cross-linked system in the NPT ensemble for 100 ns. After the system is
  cross-linked, the distance between pores shrinks by 0.4 \AA~and the distance
  between layers increases by 0.04 \AA. All major features are still present in
  the simulated XRD patterns, however at lower intensities
  (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using the
  Nernst-Einstein relation and found that it is lower in the crosslinked system
  (Fig.~\ref{fig:IC_xlink}).

  %BJC: add units
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{rzplot_xlink.png}
	\caption{}~\label{fig:rzplot_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{IC_xlink.png}
	\caption{}~\label{fig:IC_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{dbwl_xlink.png}
	\caption{}~\label{fig:dbwl_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{p2p_xlink.png}
	\caption{}~\label{fig:p2p_xlink}
  \end{subfigure}
  \caption{(a) Reflections produced by the crosslinked configuration fade
	  relative to the uncrosslinked system. The colorbar shown is the same used for
	  the uncrosslinked system. (b) The ionic conductivity is smaller relative to the
	  uncrosslinked system, but still much larger than the experimental value. When
	  the system is cross-linked, The distance between layers increases (c) and the
	  pore spacing decreases (d)}~\label{fig:xlink}
  \end{figure}

@@

  Finally, we verified that our conclusions do not change when the system is
  crosslinked by the algorithm we implemented. The diffraction pattern weakens
  relative to the uncrosslinked system, the ionic conductivity drops by a factor
  of ca. 1.5, in closer agreement with experiment, the pore spacing decreases and
  the membrane becomes thicker.

>>>>

  The permeability-selectivity tradeoff has the potential to be overcome by
  designing membranes at the molecular level. Next-generation nanoporous
  membranes with high selectivity, permitted by a precisely controlled pore size,
  and high permeability, allowed by its porous architecture, have the potential to
  replace traditional RO and NF membrane technologies.

@@

  Polymeric membranes based on self-assembling lyotropic liquid crystals (LLCs)
  are suitable candidates for aqueous separation applications. LLCs are
  amphiphilic molecules that share the characteristic ability of nanostructured
  porous membrane materials to create highly ordered nanostructures with the
  added benefits of low cost and synthetic techniques feasible for large-scale
  production \cite{feng_scalable_2014}. LLC systems created by the monomer
  Na-GA3C11 (Fig.~\ref{fig:monomer}) have been extensively studied experimentally
  \cite{feng_scalable_2014,smith_ordered_1997,zhou_supported_2005,resel_h2-phase_2000,feng_thin_2016}.
  The neat Na-GA3C11 monomer forms the thermotropic liquid crystalline (TLC)
  Col\textsubscript{h} phase (Fig.\ref{fig:assembly}). The presence of ca. 10
  wt\% added water results in formation of the lyotropic H\textsubscript{II}
  phase. In both cases, Na-GA3C11 monomers assembles into mesophases made of
  hexagonally packed, uniform size, cylinders with hydrophilic head groups
  oriented inward towards the cylinder center. This LLC assembly is then
  polymerized in situ to form a cross-linked polymer network to stabilize the
  structure. The hydrophilic region can act as a pore for aqueous separations
  \cite{zhou_supported_2005}.  One can envision tailoring the pore region for
  specific separations by changing the monomer chemistry
  \cite{resel_h2-phase_2000}.

  Research into polymerized LLC membranes has been revived in recent years.
  During early stages of exploration, mesophases formed by Na-GA3C11 could not be
  macroscopically aligned, resulting in low flux membranes, and no clear route
  towards scalable and economical filtration. In 2014, Feng et al.~showed that
  one can align the mesophases using a magnetic field with subsequent
  cross-linking to lock the structure in place \cite{feng_scalable_2014}. In 2016,
  Feng et al. showed that one could obtain the same result using a second
  technique termed soft confinement \cite{feng_thin_2016}.

@@

  Our current understanding of LLC systems is not rich enough to be able to
  precisely design membranes for specific separations. Over the past 20 years,
  H\textsubscript{II}-phase LLC polymer membrane studies have been limited
  primarily to Na-GA3C11 with some characterization done after minor structural
  modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing
  \cite{resel_structural_2000}. In a later study of rejection performance, it
  was shown that membranes formed by cross-linked Na-GA3C11 in the
  H\textsubscript{II} phase cannot separate solutes less than 1.2 nm in
  diameter because the pores are too large \cite{zhou_supported_2005}. We do not
  yet understand how to controllably reduce the effective pore size or how to
  tune the chemical environment in the nanopores of this or related materials for
  effective water desalination and small organic molecule separations. The only
  source of predictive modeling for LLC systems have been macroscopic models that
  likely do not adequately describe transport at these length scales
  \cite{hatakeyama_water_2011}. It will be challenging to efficiently narrow down
  the large design space in a laboratory setting without a robust model.

@@

  A simple molecular simulation study of a similar molecule suggested that
  there are 4 monomers in each layer \cite{zhu_methacrylated_2006}. Their 
  model likely does not sufficiently model the chemical environment present
  in the real system since it only contains 16 total monomers.
%BJC3: rephrased
%  However, their estimate is based on a simulated system that contains 16 total
%  monomers which likely does not sufficiently model the chemical environment
%  present in the real system
  A separate calculation based on the estimated volume of the liquid crystal
  monomers proposes that each layer consists of seven monomers
  \cite{resel_structural_2000}. Our best chance to answer this question is by
  using a molecular model orders of magnitude larger than any other reported
  atomistic liquid crystal membrane simulation, as we present here. We will
  directly change the layer composition and note its effect on membrane
  structure.

@@

  The $\pi$-$\pi$ stacking interactions may be a driving force of self-assembly
  in this system \cite{gazit_possible_2002}. Gas phase ab initio studies of
  benzene dimers have shown a clear energetic advantage for parallel displaced
  and T-shaped $\pi$-$\pi$ stacking conformations over a sandwiched
  conformation \cite{sinnokrot_estimates_2002}. Substituted benzene rings
  exhibit an even stronger $\pi$-$\pi$ stacking attraction which favors the
  parallel displaced configuration in all cases except where the substitutions
  are extremely electron withdrawing.
  \cite{waller_hybrid_2006,ringer_effect_2006} In this study, we compare
  simulated X-ray diffraction patterns to experiment in order to deduce which
  stacking configuration is most likely.

@@

	\item Within R-alkanes, are four spots of higher relative intensity.
	Accordingly, we name these reflection R-spots. The location of all spots is $\sim 37$
	degrees from the $q_z$ axis in their respective quadrants. In many liquid
	crystal systems one can explain the spots by the tilt angle of the alkane chains
	with respect to the membrane plane \cite{govind_simple_2001}.

	\item The final feature corresponds to the spacing and symmetry of the
	d\textsubscript{100} plane. This plane is geometrically related the distance
	between pores. The feature, which we named R-pores, is characterized by dots
	along the equatorial axis defined when $q_z$ = 0. The spacing between dots is
	indicative of the hexagonal symmetry of the packed pores. We observe the same
	information with higher resolution using SAXS (Fig.~\ref{fig:SAXS}).

@@

  We generated an ensemble of characteristic, low-energy vacuum monomer
  configurations by applying a simulated annealing process to a
  parameterized monomer. We cooled monomers from 1000K to 50K over 10
  nanoseconds. We randomly pulled a low energy configuration from the
  trajectory then reassigned charges using \texttt{molcharge}. Using the new
  charges, we annealed the monomer system again and pulled a random monomer
  configuration from the trajectory which we used for full system
  construction (Figure~\ref{fig:python}a). See supplemental information for
  further detail.

@@

  We chose the layer spacing for the initial configuration based on R-$\pi$ and
  then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of the aromatic head groups would be coplanar with the xy
  plane. We explored three different initial layer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial layer spacing of 5
  \AA. We briefly explored a third system with an initial layer spacing of 10
  \AA. If we initially space layers too far apart, they will collapse on each
  other while simultaneously slipping in between layers of adjacent pores, which
  leads to an artificially thick membrane with pores spaced closely together.
  Further details of simulations with layer stacked 10 \AA apart are in the supplemental
  information.

@@

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

@@

  \subsection{Cross-linking}

  In order to fully match synthetic procedures, we created a cross-linking
  algorithm that one can apply to equilibrated structures. The purpose of
  cross-linking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data.

  We developed the algorithm based on the known reaction mechanism.
  Cross-linking of this system is a free radical polymerization (FRP) taking
  place between terminal vinyl groups present on each of the three monomer tails.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algoritm
  selects eligible bonding carbon atoms based on a distance cut-off. The topology
  is updated with new bonds and dummy hydrogen atoms are changed to appropriate
  hydrogen types.  Head-to-tail addition was the only propagation mode considered
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

  Our implementation requires long simulation times to achieve high cross-link 
  densities. A typical cross-linking procedure can take up to 24 hours. In
  order to collect equilibrated data, further NPT simulation is necessary. We
  typically run a cross-linked system for an additional 100 ns to allow the system
  to readjust. For those reasons we did not cross-link all systems tested, but only
  the most promising structure. We show that cross-linking does not significantly
  change any of our drawn conclusions in Section 3.6.

@@

  We normalized the colorbars on all diffraction patterns relative to
  R-alkanes. We believe that the alkane-alkane density, averaged over all angles,
  is the feature most likely to be replicated between experiment and simulation.
  Other features are dependent on system ordering which is likely to have some
  dependence on initial configuration. We calculated the average intensity within
  R-alkanes of the experimental pattern, $I_{avg}$. We exclude intensities
  within $\pm$ 30\degree~of the meridional axis defined by $q_r=0$ , since the simulated patterns differ
  from experiment in those regions. Specifically, in contrast to the
  experimental WAXS pattern, R-$\pi$ appearing in simulated diffraction patterns
  intersects with R-alkanes (See Fig. \ref{fig:XRDsim}). We multiplied $I_{avg}$
  by a scaling factor of 2.5. Intensities that appear in the experimental pattern
  $\geq$ 2.5$I_{avg}$ receive colorbar values of 1. We apply the same scaling
  method to the simulated patterns. We carefully chose a scaling factor of 2.5 in
  order to (1) visibly display all features in the experimental pattern and (2)
  to allow us to compare the relative intensities of features between simulated
  and experimental diffraction patterns.

@@

  We calculated ionic conductivity using two different methods for robustness.
  The Nernst-Einstein relation relates the DC ionic conductivity to ion
  diffusivity, $D$, concentration, $C$, ion charge, $q$, the Boltzmann constant,
  $k_b$, and temperature, $T$: $$\sigma = \dfrac{q^2CD}{k_b T}$$ 
  We measured sodium ion diffusion coefficients by calculating the slope
  of the linear region of the z-direction mean square displacement curve as
  indicated by the Einstein relation \cite{einstein_investigations_1956}. We
  visualized the MSD plot to determine where to begin and end a linear fit. We
  measured ion concentration with respect to the volume of the entire unit cell.

@@

%MRS4: can we come up with another name?  I still think disordered pores is a decent name, since there is no pi-stacking layering.  ``structures that form when layers are 5 AA apart'' is not so good.  or ``more disordered''?
%BJC3: agreed, I'll stick with disordered pores
%  The $z$-direction correlation functions show that layers in our model prefer
%  to stack further apart than 3.7 \angstrom~as suggested by experiment.
  When we build systems with layers stacked 5.0 \AA~apart and then let them
  equilibrate, we observe long-term stability of a qualitatively different
  configuration, suggesting that we have found another metastable free energy
  basin, further corroborating (\ref{point:metastable}). For reasons to become
  apparent later, we will refer to this class of systems as the disordered pore
  basin. Systems which we build with layers initially stacked 3.7 \AA~apart will
  be referred to as the ordered basin. We studied each basin in both the parallel
  displaced and sandwiched configurations.

@@

  In all cases, the space in the pore region is filled with sodium ions and
  head groups. In the ordered pore systems, there is less density in the center of
  the pore indicating a more ordered pore structure. In both the sandwich and
  parallel displaced configurations, we see the density of head groups and sodium
  ions fall to less than 50\% of its maximum at r = 0
  (Fig.~\ref{fig:layered_density}). The situation is most pronounced in the
  sandwiched configuration where the maximum head group density occurs 0.44 nm
  from the pore center. The parallel displaced configuration reaches its maximum
  0.35 nm from the pore center. In contrast, both disordered pore systems show
  very little difference in density from its maximum. This implies a more uniform
  distribution of head groups within the pore center. 
  % High vacancy is less likely to be stable in the real system

@@

  The system's structure and physical characteristics did not change
  significantly when we applied the cross-linking algorithm to the equilibrated
  parallel-displaced configuration built with 5 monomers per layer. We simulated
  the cross-linked system in the NPT ensemble for 100 ns. After the system is
  cross-linked, the distance between pores shrinks by 0.4 \AA~and the distance
  between layers increases by 0.04 \AA. All major features are still present in
  the simulated XRD patterns, however at lower intensities
  (Fig.~\ref{fig:rzplot_xlink}). We calculated the ionic conductivity using the
  Nernst-Einstein relation and found that it is lower in the cross-linked system
  (Fig.~\ref{fig:IC_xlink}).

  %BJC: add units
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{rzplot_xlink.png}
	\caption{}~\label{fig:rzplot_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{IC_xlink.png}
	\caption{}~\label{fig:IC_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{dbwl_xlink.png}
	\caption{}~\label{fig:dbwl_xlink}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
	\centering
	\includegraphics[width=\textwidth]{p2p_xlink.png}
	\caption{}~\label{fig:p2p_xlink}
  \end{subfigure}
  \caption{(a) Reflections produced by the cross-linked configuration fade
	  relative to the uncross-linked system. The colorbar shown is the same used for
	  the uncross-linked system. (b) The ionic conductivity is smaller relative to the
	  uncross-linked system, but still much larger than the experimental value. When
	  the system is cross-linked, The distance between layers increases (c) and the
	  pore spacing decreases (d)}~\label{fig:xlink}
  \end{figure}

@@

  Finally, we verified that our conclusions do not change when the system is
  cross-linked by the algorithm we implemented. The diffraction pattern weakens
  relative to the uncross-linked system, the ionic conductivity drops by a factor
  of ca. 1.5, in closer agreement with experiment, the pore spacing decreases and
  the membrane becomes thicker.
